%!TEX root = userguide.tex
\section{Introduction}


\subsection{What is PEBBL?}

PEBBL (\emph{P}arallel \emph{E}numeration and
\emph{Branch}-and-\emph{B}ound \emph{L}ibrary) is a C++ class library
for constructing serial and parallel branch-and-bound optimization
algorithms.  It is a \emph{framework}, \emph{shell}, or
\emph{skeleton} that handles the generic aspects of branch and bound,
allowing the developer to focus primarily on the unique aspects of
their particular branch-and-bound algorithm.  
It is thus similar to other software projects such
as PUBB~\cite{SHH95,SHH97}, BoB~\cite{bob95}, PPBB-Lib~\cite{PPBB96},
and ALPS~\cite{RLS04} (PEBBL's development has significantly
influenced the architecture of ALPS).  PEBBL has a number of unique
features, including very flexible parallelization strategies and the
ability to enumerate near-optimal solutions.

In principle, one can build an arbitrary branch-and-bound method atop
PEBBL by defining a relatively small number of abstract methods.  By
defining a few more methods, the algorithm can be immediately
parallelized.  PEBBL contains numerous tuning parameters that can
adapt the resulting parallel implementation to any parallel
architecture that supports an MPI message passing
library~\cite{SOHWD96}.


\subsection{Changes in this release}
The principal changes in PEBBL since the prior release (version 1.6) are:
\begin{enumerate}
\item The ability to run within a user-specified MPI communicator, as opposed
to only within \texttt{MPI\_COM\_WORLD}.
%%% Uncomment when we actually implement this...
% \item Support for multi-level parallelism throughout the search process ---
% for example having clusters of 100 MPI processes to process each subproblem,
% but using many such clusters in parallel to explore different parts of the
% search tree simultaneously \textbf{Still have to do this!}
\item Distribution through \texttt{github.com} using \texttt{git}, rather than
through a Sandia labs server using subversion (\texttt{svn}).
\item Adoption of the \texttt{cmake} build system in place of \texttt{autoconf}.
\item Separation from the ACRO family of packages.  PEBBL now consists of just
one package and has no external dependencies.  Some parts of the old UTILIB
package are now embedded with PEBBL.
\item As part of the separation from ACRO, the directory structure of PEBBL
has been completely reworked and greatly simplified, as described in
Section~\ref{sec:dirstruct} below.  Related changes include
\begin{itemize}
\item C++ \texttt{\#include} directives for PEBBL are now naturally
self-documenting in that header file names starting with \texttt{pebbl/}.
There is no longer a ``flattened'' copy of the include files structure, so
subdirectories such as \texttt{pebbl/bb} or \texttt{pebbl/pbb} must be
explicitly specified.
\item The file names for the included Python scripts now end in \texttt{.py}.
\end{itemize}
\item Miscellaneous bug fixes.
\end{enumerate}


\subsection{The Genealogy of PEBBL}

Most of the development work on PEBBL has been carried out by
\begin{itemize}
\item Jonathan Eckstein, Rutgers University
\item William Hart, Sandia National Laboratories
\item Cynthia A. Phillips, Sandia National Laboratories.
\end{itemize}
PEBBL was originally the ``core''
layer of the PICO (\emph{P}arallel \emph{I}nteger and
\emph{C}ombina\-torial \emph{O}ptimization) package.  PICO was designed
to solve mixed integer programming problems, but included a ``core''
layer supporting implementation of arbitrary branch and bound
algorithms.  In the Spring of 2006, the development team decided to
distribute this core layer as a software package in its own right,
changing its name from ``the PICO core'' to PEBBL.  Much of PEBBL's
basic design is thus described in preliminary
publications concerning PICO~\cite{EHP97,EPH00,EPH00a}.  In fact,
significant portions of this user guide are derived
from Eckstein \emph{et al.}~\cite{EPH00a,EPH00}.

PEBBL's parallelization strategies are partially patterned after
CMMIP~\cite{Eck94,Eck97}, a parallel mixed integer programming solver
developed for the Thinking Machines CM-5 parallel supercomputer.
CMMIP, however, was specifically designed for mixed integer
programming, and to take advantage of particular features of the CM-5
architecture.  PEBBL is more generic in two senses: it is a framework that
one can use to implement any branch-and-bound algorithm, and it is
designed to run in a generic message-passing environment.  The ABACUS
package~\cite{JT98} also influenced some of PEBBL's design.
